My title

### 2024-11-29 paper/paper.qmd
generated new pdf
fixed typos

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Hyunje Park
thanks: "Code and data are available at: [https://github.com/davidpxrk/childrens-book-rating](https://github.com/davidpxrk/childrens-book-rating). Special thanks to Rohan Alexander for his help!"
date: today
date-format: long
abstract: "This paper examines the rating of childrens' books using the . By using a Bayesian logistic regression and data from the Alex Cookson (CITATION), I analyzed the relationship between various characteristics of a childrens' book, and whether or not it influenced if the book had a high rating or not. The insight from this research not only clarify what makes a high rated childrens' book so high rated, but also enhance future authors to push the boundaries of their literature."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(palmerpenguins)
library(kableExtra)
library(arrow)
library(readr)
library(rstanarm)
library(modelsummary)
library(dplyr)

library(tidyverse)
library(here)
library(rstanarm)
library(modelsummary)
library(ggplot2)
library(knitr)

@@

For generations, children's books have been a catalyst in children's development. Through captivating fairytales and fables of stories about adventures, heroes, magical forests and magic, children gain experiences on feelings and thoughts, learning to cope with inhibitions, vulnerability and shyness (CITATION). Beyond educational purposes, children's literature can positively influence mental wellbeing, feelings and behavior. Given these significant developmental benefits, it can be said that the quality of the children's book matters greatly; choosing a well-written book can affect how it nurtures the next generation of mature, emotionally resilient individuals. This is why book rating systems hold significant importance; allowing readers to rate books on a scale from 0-5 scale (most commonly used scale) helps parents and educators assess whether a book is worth giving to children.

In this 0-5 rating system, a score of 4 or above is often seen as the benchmark for a "highly rated" book. This is influenced by central tendency bias, where people naturally gravitate towards a moderate score, avoiding extremes like 0 or 5 to appear more balanced and objective. A rating such as 4, in particular, suggests a strong endorsement without overstepping into exaggeration. Given this, a critical question arises; What factors of a book contribute to the likelihood of the book being "highly-rated" (a score above 4)?

In this paper, I analyzed how the characteristics of a book, such as book type, page count, publishing year, and rating counts affected the likelihood of a book being "highly-rated" using the Children's Book Ratings Data (CITATION). First, after data cleaning, I selected 9 variables on children book characteristics for my analysis in (SECTION 2). Then, a logistic regression model was created to predict the probability of the book being "highly-rated", based on the chosen book characteristic variables.

@@

The estimand of this paper is the probability that a book is highly-rated (has an average rating of over 4 on a 0-5 scale), based on book characteristics. It is difficult to measure the exact number as there are millions of children's books that are published and not all of them will be accessed due to various issues. For examples, children's books from different countries may have different ratings. Therefore in this paper, we attempt to estimate the estimand using a logistic regression model which is fitted using a sample from the Children's Book Rating dataset (CITATION)

@@

The dataset used in this paper was obtained from COOKSON (CITATION), who sourced it from Goodreads (CITATION). Cookson's dataset contained over 9,000+ records of children's books that contained information on the book such as title, author, ratings, publisher and more. In the initial step, data-cleaning was performed and relevant variables were selected. First, entries with missing information were filtered out, and variables that were irrelevant to this study was also removed, which included variables such as book title. After cleaning, there were 9240 entries in the cleaned dataset.

@@

data
# TABLE 1: CLEANED DATASET
first_half <- data %>% select(1:6)

@@

Table 1 (CROSS REFERENCE) presents the first 5 rows from the cleaned dataset, that contains information on Children's book.

@@

- `rating_count`: An integer that represents how many ratings the book received

- `publish_year`: An integer that represents the year the book was published

- `cover`: A categorical variable that represents the type of the book, which can include:
  - `Ebook`: The book is in a digital format.
  - `Kindle`: The book is only available through Kindle e-reader tablet.
  - `Board book`: The book is a picture book designed for the youngest of children, babies, infants.
  - `Hardcover`: The book has a rigid protective cover.
  - `Paperback`: The book is has a thick paper cover.

Table 2 (CITATION) shows a summary of detailed statistics about the cleaned dataset. It showed that books that were considered "high rated" had more pages on average than low rated books. Furthermore, high rated books were on average published 2 years before low rated books, around the year 2003. In addition to this, the amount of ratings high-rated books received far exceed low rated books, at about 4x. Kindle covers had the highest proportion of "high-rated" books out of any other cover types. These variables are key in the analysis, and are further shown in (CITE HERE)

```{r}
#| label: summary_statistic_cleaned_dataset
#| fig-cap: "Summary Statistic of the Cleaned Dataset"
#| echo: false

summary_table <- data %>%
  filter(!is.na(rated_high)) %>%  
  group_by(rated_high) %>%       
  summarise(
    avg_pages = mean(pages, na.rm = TRUE),          
    avg_publish_year = mean(publish_year, na.rm = TRUE),
    avg_rating_count = mean(rating_count, na.rm = TRUE),
    highest_proportion_cover = data %>%
      group_by(cover) %>%
      summarise(high_rated_pct = mean(rated_high, na.rm = TRUE)) %>%
      slice_max(high_rated_pct, n = 1) %>%
      pull(cover)
  )

@@

```{r}
#| label: fig-most-common-covers
#| fig-cap: "Summary Statistic of the Cleaned Dataset"
#| echo: false

data2 <- data %>%filter(rating >= 4)

cover_proportions <- data %>%
  group_by(cover) %>%
  summarise(high_rated_pct = mean(rated_high, na.rm = TRUE)) %>%
  arrange(desc(high_rated_pct))  # Sort by highest proportion

# Create the horizontal bar plot
ggplot(cover_proportions, aes(x = high_rated_pct, y = reorder(cover, high_rated_pct), fill = high_rated_pct)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  labs(
    title = "Proportion of High-Rated Books by Cover Type",
    x = "Proportion of High-Rated Books",
    y = "Cover Type"
  ) +
  scale_x_continuous(labels = scales::percent_format()) +  # Format x-axis as percentages
  theme_minimal() +
  theme(
    axis.text = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, size = 14)
  )

@@

(CITATION HERE PICTURE ABOVE) shows the number of cover types for "high-rated" books (rated_high = 1). It is evident that Paperback and Hardcover books make up the majority of "high-rated" books, which could potentially reflect the aesthetic and collectible value of Paperback/Hardcover books over something non-physical like an Ebook. This illustrates that physical books could be more appealing for readers due to ownership bias, greatly increasing their rating. This finding also reflects the result from (TABLE)

```{r}
#| label: fig-page-count
#| fig-cap: "Summary Statistic of the Cleaned Dataset"
#| echo: false
avg_pages <- aggregate(pages ~ rated_high, data = data, FUN = mean, na.rm = TRUE)

ggplot(avg_pages, aes(x = factor(rated_high), y = pages, fill = factor(rated_high))) +
  geom_bar(stat = "identity") +
  labs(
    title = "Average Pages by High Rated",
    x = "High Rated (1 = Yes, 0 = No)",
    y = "Average Pages",
    fill = "High Rated"
  ) +
  theme_minimal()

(CTATION HERE) shows that generally, "high-rated" books have more page count than low-rated books. This suggests that higher page counts could correlate with higher ratings as more pages could allow deeper storytelling, richer world-building or character development, which can result in a more immersive experience for the reader. This graph illustrates the importance for authors to use more pages for properly structured narratives.

```{r}
#| label: fig-publish-year
#| fig-cap: "Summary Statistic of the Cleaned Dataset"
#| echo: false

data2 <- data %>% filter(rating >= 4)
data3 <- data %>% filter(rating < 4)
data2 <- data2 %>% mutate(group = "High Rating (>= 4)")
data3 <- data3 %>% mutate(group = "Low Rating (< 4)")

@@

# Create a combined scatter plot with facets
ggplot(combined_data, aes(x = publish_year, y = rating)) +
  geom_point() +
  theme_minimal() +
  facet_wrap(~group, ncol = 1, scales = "free_y") +
  geom_vline(data = avg_publish_years, aes(xintercept = avg_publish_year), 
             color = "red", linetype = "dashed", size = 1) 
```
(CITATION) shows the relationship between year of publish and the book's rating. Where the horizontal axis shows the year of publishment, and vertical axis shows the rating of the book. By looking at the red line of fit (average year of publishment), "high-rated" books tend to be published before low-rated books. This could be due to selective survival where older books that remain in circulation could be considered "classics" and have stood the test of time.This is further discussed in (DISCUSSION HERE)

```{r}
#| label: fig-rating-count
#| fig-cap: "Summary Statistic of the Cleaned Dataset"
#| echo: false
#| 
data2 <- data %>% filter(rating >= 4)

ggplot(data, aes(x = rating, y = rating_count)) +
  geom_point() 
```

@@

Furthermore, the irrelevance of the number of ratings a book receives could be explained by herd effect; where readers rely on the ratings and reviews of others to make their own rating decisions. Therefore, once a book has many reviews, its average rating stabilizes, making the number of ratings irrelevant to the overall evaluation.

@@

The results in SECTION (RESULTS), indicated a big positive relationship between book cover types and its "highly-rated" status. In particular, the "Kindle Edition" book cover had the highest coefficient (NUMBER HERE). A potential explanation could be that ebooks (such as Kindle Edition) are often preferred over physical books for their convenience, affordability and accessibility. Additionally ebooks are often cheaper due to saving material costs and saving physical space. In Middle-Eastern countries, 74.6% students preferred e-books in terms of easy to carry and 80.6% of them spent more time reading from e-books than printed books (CITATION HERE). Thus, it could appear that e-books have an advantage in attracting those seeking convenience and personalization in their reading, potentially explaining its high-rating over any other book cover types.

@@

The second explanation could be that older books may benefit from nostalgia, where readers could assign higher ratings due to a stronger emotional connection and cultural significance they hold. While this

@@

We used a posterior predictive check in FIGURE (CITE) to evaluate how well the model fits the data. This compares the observed data Y (CHANGE) against the replicated data y-REP (CHANGE) which was generated by our model in SECTION 3 (CITE). The poterior distribution fitting perfectly suggests that the Bayesian logistic regression model is a good model fit.

@@

# delete this later
model_file <- readRDS(file = here::here("models/model.rds"))

pp_check(model_file) +
  theme(legend.position = "bottom")

@@

The trace plot shows osciliating horizontal lines, with overlaps between chains, showing no signs of issues with the model.

@@

>>>>

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Hyunje Park
thanks: "Code and data are available at: [https://github.com/davidpxrk/childrens-book-rating](https://github.com/davidpxrk/childrens-book-rating). Special thanks to Rohan Alexander for his help!"
date: today
date-format: long
abstract: "This paper examines the factors influencing the rating of children's books using the data from Alex Cookson (CITATION). By using a Bayesian logistic regression, I analyzed how various book characteristics impact the likelihood of a book receiving a high rating. The results showed that the book's cover type had the biggest positive impact on a book's rating. The insight from this research not only clarify the key elements that make high rated childrens' book so high rated, but also enhance future authors to push the boundaries of their literature and writing choices."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(palmerpenguins)
library(kableExtra)
library(arrow)
library(readr)
library(rstanarm)
library(modelsummary)
library(dplyr)
library(here)
library(rstanarm)
library(modelsummary)
library(ggplot2)
library(knitr)
library(scales)

@@

For generations, children's books have been a catalyst in children's development. Through captivating fairytales and fables of stories about adventures, heroes, magical forests and magic, children gain experiences on feelings and thoughts, learning to cope with inhibitions, vulnerability and shyness (CITATION). Given these significant developmental benefits, it can be said that the quality of the children's book matters greatly; choosing a well-written book can affect how it nurtures the next generation of mature, emotionally resilient individuals. This is why book rating systems hold significant importance; allowing readers to rate books on a scale from 0-5 (most commonly used scale) helps parents and educators assess whether a book is worth giving to children.

In this 0-5 rating system, a score of 4 or above is often seen as the benchmark for a "highly rated" book. This is influenced by central tendency bias, where people naturally gravitate towards a moderate score, avoiding extremes like 0 or 5 to appear more balanced and objective. A rating such as 4, in particular, suggests a strong endorsement without overstepping into exaggeration. Given this, a critical question arises; What factors of a book contribute to the likelihood of the book being categorized as "highly-rated" (a score above 4)?

In this paper, I analyzed how the characteristics of a book, such as cover type, page count, publish year, and rating counts affected the likelihood of a book being "highly-rated" using the Children's Book Ratings Data (CITATION). First, after data cleaning, I selected 9 variables on children book characteristics for my analysis in (SECTION 2). Then, a logistic regression model was created to predict the probability of the book being "highly-rated", based on the chosen book characteristic variables.

@@

The estimand of this paper is the probability that a children's book is "highly-rated" (has a rating over 4 on a 0-5 scale), based on book characteristics. It is difficult to measure the exact number as there are millions of children's books that are published and not all of them will be accessed. For examples, children's books may have variations in ratings across countries. Therefore in this paper, we attempt to estimate the estimand using a logistic regression model which is fitted using a sample from the Children's Book Rating dataset (CITATION).

@@

The dataset used in this paper was obtained from COOKSON (CITATION), who sourced it from Goodreads (CITATION). Cookson's dataset contained over 9,000 records of children's books that contained information of books such as title, author, ratings, publisher and more. In the initial step, data-cleaning was performed and relevant variables were selected. First, entries with missing information were filtered out, and variables that were irrelevant to this study were also removed, which included variables such as book title. Then, a new variable called "highly-rated" was created, which was a binary variable indicating if the book had a rating over 4 (1 if it exceeded 4, 0 else). After cleaning and creating this new variable, there were 9240 entries in the cleaned dataset.

@@

# TABLE 1: CLEANED DATASET
first_half <- data %>% select(1:6)

@@

Table 1 (CROSS REFERENCE) presents the first 5 rows from the cleaned dataset, which contains information on Children's book.

@@

- `rating_count`: An integer representing the number of ratings the book received

- `publish_year`: An integer representing the year the book was published

- `cover`: A categorical variable representing the type of the book, which can include:
  - `Ebook`: The book is in a digital format.
  - `Kindle`: The book is only available through Kindle e-reader tablet.
  - `Board book`: The book is a picture book designed for the youngest of children, babies, infants.
  - `Hardcover`: The book has a rigid protective cover.
  - `Paperback`: The book is has a thick paper cover.

Table 2 (CITATION) shows a summary of detailed statistics about the cleaned dataset. It showed that books that were considered "high rated" had more pages on average than low-rated books. Furthermore, high rated books were on average published 2 years before low rated books, with publish year of 2005 and 2003 for the low-rated books and high-rated books respectively. High-rated books also received far more ratings on average at 8000 ratings, more than four times of low-rated books at an average of 2000 reviews per book. Lastly, Kindle covers had the highest proportion of "high-rated" books out of any other cover types (as well as low-rated books). These variables are key in the analysis, and are further mentioned in (CITE HERE)

```{r}
#| label: summary_statistic_dataset
#| fig-cap: "Summary Statistic of the Cleaned Dataset"
#| echo: false
#| warning: false
summary_table <- data %>%
  filter(!is.na(rated_high)) %>%  
  group_by(rated_high) %>%       
  summarise(
    avg_pages = mean(pages, na.rm = TRUE),          
    avg_publish_year = mean(publish_year, na.rm = TRUE),
    avg_rating_count = mean(rating_count, na.rm = TRUE),
    highest_proportion_cover = data %>%
      group_by(cover) %>%
      summarise(high_rated_pct = mean(rated_high, na.rm = TRUE)) %>%
      slice_max(high_rated_pct, n = 1) %>%
      pull(cover)
  )

@@

```{r}
#| label: fig-most-common-covers
#| fig-cap: "Summary Statistic of the Cleaned Dataset"
#| echo: false
#| warning: false

data3 <- data %>%
  filter(!is.na(cover) & !is.na(rated_high))

cover_proportions <- data3 %>%
  group_by(cover) %>%
  summarise(
    total_books = n(),                                # Total number of books for each cover type
    high_rated_books = sum(rated_high, na.rm = TRUE), # Total number of highly-rated books for each cover type
    high_rated_pct = high_rated_books / total_books   # Proportion of highly-rated books
  ) %>%
  arrange(desc(high_rated_pct))  # Sort by the highest proportion

# Create the bar plot
ggplot(cover_proportions, aes(x = high_rated_pct, y = reorder(cover, high_rated_pct), fill = high_rated_pct)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  geom_text(aes(label = percent(high_rated_pct, accuracy = 0.1)), 
            hjust = -0.1, 
            size = 3.5) +  # Add percentage labels
  labs(
    title = "Proportion of Highly-Rated Books by Cover Type",
    x = "Proportion of Highly-Rated Books",
    y = "Cover Type"
  ) +
  scale_x_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) + # Format x-axis as percentages
  theme_minimal() +
  theme(
    axis.text = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, size = 14),
    plot.margin = margin(10, 50, 10, 10)
  )

@@

(CITATION HERE PICTURE ABOVE) shows the proportion of "highly-rated" books for each cover type. `Kindle Edition` books had the highest proportions of highly-rated books at 77.8%, followed by Paperback and Ebook, at 64.7% and 61.3% respectively. This illustrates that generally, digitally formatted books had higher proportions of "highly-rated" books compared to physical formats. This finding also reflects the result from (TABLE)

```{r}
#| label: fig-page-count
#| fig-cap: "Summary Statistic of the Cleaned Dataset"
#| echo: false
#| warning: false

avg_pages <- aggregate(pages ~ rated_high, data = data, FUN = mean, na.rm = TRUE)

# Create the bar plot with text labels
ggplot(avg_pages, aes(x = factor(rated_high), y = pages, fill = factor(rated_high))) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(pages, 1)), vjust = -0.5, size = 4) +  # Add labels above the bars
  labs(
    title = "Average Pages by High Rated",
    x = "High Rated (1 = Yes, 0 = No)",
    y = "Average Pages",
    fill = "High Rated"
  ) +
  theme_minimal()
```

(CTATION HERE) shows that "high-rated" books have more page count than low-rated books. Highly-rated books had an average page count of 66.1, compared to the 57.6 average page count of low-rated books.

```{r}
#| label: fig-publish-year
#| fig-cap: "Summary Statistic of the Cleaned Dataset"
#| echo: false
#| warning: false

data2 <- data %>% filter(rating >= 4)
data3 <- data %>% filter(rating < 4)
data2 <- data2 %>% mutate(group = "High Rated Books (>= 4)")
data3 <- data3 %>% mutate(group = "Low Rated Books (> 4)")

@@

# Create a combined scatter plot with facets
ggplot(combined_data, aes(x = publish_year, y = rating)) +
  geom_point() +
  theme_minimal() +
  facet_wrap(~group, ncol = 1, scales = "free_y") +
  geom_vline(data = avg_publish_years, aes(xintercept = avg_publish_year), 
             color = "red", linetype = "dashed", size = 1) 
```
(CITATION) shows the relationship between a book's rating and the publish year. The horizontal axis shows the publish year, and vertical axis shows the rating of the book. The graph is split into two vertically, the top plot showing ratings of "high-rated" books and low-rated books. Low-rated books showed a higher concentration of publish year towards the right, suggesting a higher average publish year. By looking at the red line of fit (average publish year), "high-rated" books tend to be published before low-rated books.

```{r}
#| label: fig-rating-count
#| fig-cap: "Summary Statistic of the Cleaned Dataset"
#| echo: false
#| warning: false

data2 <- data %>% filter(rating >= 4)

ggplot(data, aes(x = rating, y = rating_count)) +
  geom_vline(xintercept = 4, color = "red", linetype = "dashed", size = 1) +
  geom_point() +
  theme_minimal()
```

@@

Furthermore, the irrelevance of the number of ratings a book receives could be explained by herd effect; where readers rely on the ratings and reviews of others to make their own rating decisions (CITATION). Therefore, once a book has many reviews, its average rating stabilizes, making the number of ratings irrelevant to the overall evaluation.

@@

The results in SECTION (RESULTS), indicated a big positive relationship between book cover types and its "highly-rated" status. In particular, the "Kindle Edition" book cover had the highest coefficient (NUMBER HERE). A potential explanation could be that ebooks (such as Kindle Edition) are often preferred over physical books for their convenience, affordability and accessibility. Additionally ebooks are often cheaper due to saving material costs and saving physical space. In Middle-Eastern countries, 74.6% students preferred e-books in terms of easy to carry and 80.6% of them spent more time reading from e-books than printed books (CITATION HERE). Thus, it could appear that e-books have an advantage in attracting those seeking convenience and personalization in their reading, potentially explaining its big relationship compared to any other book cover types.

@@

The second explanation could be that older books may benefit from nostalgia, where readers could assign higher ratings due to a stronger emotional connection and cultural significance they hold.

## Weaknesses and Next Steps

Regional differences were not considered in this paper. A book that appeals a reader in a country might not have the same appeal in another due to the differences in cultural context, historical experiences or social norms. Furthermore, regional differences in terms of accessibility of books (such as quality of translated versions or availability) could also impact ratings.

In terms of the dataset, the number of entries were relatively low (a little over 9,000 entries). This limited number of entries could have affected the accuracy of the results, where certain factors such as book cover types might have been significantly overrepresented or underrepresented, skewing the analysis.

Future research should aim to collect more representative data, potentially including datasets from diverse sources. Additionally, regional factors related to the book should be investigated. By exploring these regional influences, we can get an broader insight as to how cultural context, historical experience and social norms play a role in book ratings. This can help make improvements for a broader international appeal.

@@

We used a posterior predictive check in FIGURE (CITE) to evaluate how well the model fits the data. This compares the observed data Y (CHANGE) against the replicated data y-REP (CHANGE) which was generated by our model in SECTION 3 (CITE). The posterior distribution fitting perfectly suggests that the Bayesian logistic regression model is a good model fit.

@@

pp_check(model_file) +
  theme(legend.position = "bottom")
```

@@

The trace plot shows oscillating horizontal lines, with overlaps between chains, showing no signs of issues with the model.

@@

```{r}
#| label: summary_statistic_cleaned_dataset72
#| fig-cap: "Summary Statistic of the Cleaned Dataset"
#| echo: false
